### 2019-10-29 Paper_shortened_pdf.Rmd
Fix spelling

abstract: |
  The question whether fathers and mothers play similar roles in children’s
  emotional adjustment has been approached differently in different research
  traditions: Developmental psychologists have focused on qualitative parenting,
  and sociologists have focused on quantitative child care involvement. This
  longitudinal study (N = 1884, age 6-12) bridges these approaches, 
  investigating qualitative and quantitative involvement, and their interaction,
  in relation to emotional adjustment. On average, evidence for mothers'
  and fathers' unique roles was limited.
  When both parents were more involved, both engaged
  more in desirable qualitative parenting. When mothers were disproportionally
  involved, however, parenting roles were more differentiated:
  Fathers displayed less
  desirable qualitative parenting, and children were less well-adjusted.
  This suggests that all family members benefit from a more equal
  childcare distribution.

@@

\vspace{\baselineskip}
Emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001]. It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013]. Emotional adjustment development is a domain in which fathers might play a unique role, because fathers encourage children to challenge themselves, in play and in life [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000]. The attachment relationships between children and parents are grounded in a balance between security and exploration [@grossmannDeveloppementAttachementAdaptation1998]. Fatherhood researchers have posited that warmth and security are more central to mother-child attachment relationships, whereas father-child relationships have been construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013]. By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities.

@@

The past decades have seen major strides in the consolidation of theories to guide fatherhood research. Inspired by increasing divorce rates and maternal employment, early fatherhood research primarily operationalized involvement in *quantitative* terms: Presence versus absence in divorced families, and the sharing of childcare duties in dual-earner families [@lambBiosocialPerspectivePaternal1987]. Over time, the notion of the father as a coparent gained ground [@pleckFatherhoodIdealsUnited1997], according to which fathers should share childrearing responsibilities with mothers, adopting roles historically fulfilled by mothers. To keep pace with changing fatherhood norms, definitions of father involvement evolved to incorporate *qualitative* dimensions, such as warmth and control [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].

Although it is now widely acknowledged that involvement quality and quantity are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely investigated together. Historically, developmental psychology has emphasized how qualitative parenting behaviors shape child development, whereas sociology and gender studies have focused on involvement quantity and its implications for maternal well-being and gender equality [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014]. The present study attempts to bridge these approaches by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.

@@

The literature linking demandingness and emotional adjustment is less conclusive [@morrisRoleFamilyContext2007]. Demandingness encompasses several parenting practices related to setting and enforcing expectations for children's behavior. Some of these might help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013]. We examined two aspects of demandingness: Consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments. Consistency has been found to correlate strongly with emotional adjustment [@brandPerceivedParentalInconsistency1990]. Inductive reasoning is thought to promote the internalization of parental expectations, including display rules for emotions, and might thus play a role in emotional adjustment development as well [@grusecImpactParentalDiscipline1994].

@@

With regard to differential involvement, we might also expect the more involved parent to display more desirable qualitative parenting behaviors, due to the effects of experience [@sabattiniRelationMothersFathers2004]. However, the benefit of parents monitoring each other's childrearing should be diminished when parents are not equally involved [@mclanahanConsequencesFatherAbsence2014]. Indeed, retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004]. Additionally, greater differential involvement might undermine both parents' involvement quality, through its effect on the marital relationship. Childcare responsibilities are a heavy burden [@deutschPaternalParticipationChild2001]. An unfair distribution strains the interparental relationship, interferes with sensitive parenting, and undermines child wellbeing [@davies2006interparental]. When care is distributed relatively equally, mothers tend to be more satisfied [@pharesFathersMothersInvolvement2009], and the family system more harmonious. Consequently, balanced co-parenting is associated with greater parental warmth [@barnettInterdependenceParentingMothers2008]. We might thus expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment.

@@

Bioecological theories of development indicate that children require a certain amount of care, and who provides this care is, to some extent, flexible [see @bronfenbrennerEcologyHumanDevelopment1979]. This suggests that the total amount of care provided should be positively associated with children's emotional adjustment. Differential involvement, on the other hand, might be negatively linked to emotional adjustment: When children perceive that one of their parents is relatively uninvolved, they likely feel less supported than when both parents are equally invested in their wellbeing [@deutschPaternalParticipationChild2001].

@@

Interactions between involvement quality and quantity have largely been overlooked in theoretical models of father involvement; perhaps in part because these two dimensions have rarely been jointly investigated [@jaffeeLifeFatherBenefits2003]. In this respect, theories of father involvement can be complemented by Feinberg's coparenting model [@feinbergInternalStructureEcological2003]. Coparenting refers to the ways parents relate to each other in the role of parent. Although the coparenting literature is typically invoked to describe childrearing in divorced families, it also holds insights for intact families. According to Feinberg, the "division of labor" in childcare, analogous to our operationalization of differential involvement, affects qualitative childrearing behaviors and children's adjustment directly, and moderates the link between these two constructs.

@@

Emotional adjusmtment was measured using the reverse-coded emotional distress subscale of the teacher-reported Strengths and Difficulties Questionnaire, on a three-point scale from "not true" to "certainly true" [e.g., "Often unhappy, downhearted or tearful"; @goodmanPsychometricPropertiesStrengths2001].

@@

We calculated mean scores for all of the study variables. Scale descriptive statistics are provided in Supplementary Table 1, and scale intercorrelations are provided in Supplementary Table 2. We deemed it defensible to assume measurement invariance, as the difference in fit between structurally invariant models, and metrically invariant models, was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).

To address our research questions, we used multilevel, autoregressive structural equation modelling [@asparouhovDynamicStructuralEquation2018]. In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean). This has several advantages: First, controlling for stable between-family differences allows us to estimate the direction of within-family effects more accurately [@hamakerCritiqueCrosslaggedPanel2015]. Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts. Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects. In this way, we included total joint childcare, the childcare distribution, and child sex as moderators.

@@

With regard to child effects, in line with hypotheses, the intercept of the cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabcrosspath)). Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more anger and inductive reasoning, but also more warmth, than usual in the next year. Contrary to hypotheses, the intercept of the effect of emotional adjustment on maternal consistency was non-significant. As hypothesized, the intercept for maternal inductive reasoning was positive. Finally, counter to hypotheses, the intercept for maternal anger was positive, not negative. In partial support of hypotheses, the intercept for the child effect of adjustment on paternal warmth was positive and significant. None of the other intercepts for child effects on paternal parenting were significant. The variances of all child effects were significant, indicating that the effect of children's emotional adjustment on their parents' childrearing behaviors differs between families.

@@

When both parents were more involved, they displayed greater warmth, consistency, and inductive reasoning. Fathers, in particular, reported substantially greater parental warmth when both parents were more involved. We found that, when mothers performed more care tasks than fathers, mothers displayed more warmth, and fathers displayed less warmth, consistency, and inductive reasoning. This increase in maternal warmth might reflect an attempt to compensate for fathers' absence [@shinnFatherAbsenceChildren1978]. These results are in line with our hypothesis that greater involvement would be associated with more desireable parenting behaviors [@sabattiniRelationMothersFathers2004]. Additionally, they suggest that, when childcare is unequally distributed, fathers engage in less desirable qualitative parenting behaviors [@craigHowMothersFathers2011].

@@

Our operationalization of quantitative involvement, in terms of total joint care and the childcare distribution, allowed us to addresses the claim that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001]. However, counter to our hypothesis, joint care was *negatively* associated with between-family differences in emotional adjustment. One potential explanation is suggested by self-determination theory: Excessive levels of involvement might undermine children's autonomy needs, thereby leading to increased emotional maladjustment [@vandergiessenPerceivedAutonomySupport2014]. Research on "helicopter parenting" supports this interpretation, as psychological wellbeing was found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014]. Another potential explanation is that the amount of quantitative care provided is not only driven by parents' motivation and ability to be involved in their children's lives, but also by children's objective needs. Factors related to children's objective needs - such as developmental delay, or atypical development - might thus drive an increase in parents' quantitative involvement, and might also be associated with children's emotional maladjustment, thereby resulting in a negative correlation between the two.

In line with predictions derived from coparenting theory, we found that children's emotional adjustment was lower in families where mothers were more involved than fathers - even after controlling for the total amount of care provided. Although substantial empirical work has indicated that the marital relationship suffers when childcare is unequally distributed [@shockleyItNotWhat2018], our research is one of few to examine the implications for children's adjustment [see also @deutschPaternalParticipationChild2001]. These correlational findings at the between-family level should not be interpreted causally, however. Future research should examine potential pathways, such as whether children's adjustment suffers when they feel like both parents are not equally invested in their wellbeing, or whether fathers are more likely to withdraw from childcare responsibilities when children display poor emotional adjustment, or whether differential involvement strains the marital relationship, and this strain spills over to children's emotional adjustment.

@@

Several strengths allow us to make unique contributions to the literature. First, mothers and fathers both completed identical parenting measures, which allowed us to disentangle and compare their unique roles in children's emotional adjustment development. Second, we incorporated both quantitative and qualitative involvement. This allowed us to provide a more nuanced overview of the links between parental involvement and emotional adjustment development than prior work, including how the distribution of quantitative childcare relates to parenting quality and emotional adjustment, and moderates links between these two constructs. Although some prior studies have also incorporated both quality and quantity of parenting, this work did not disentangle the total amount of care children received from the relative distribution of care tasks [see @deutschPaternalParticipationChild2001]. We addressed this limitation by re-expressing the variance of maternal and paternal quantitative involvement into joint total involvement, and differential involvement. Third, we used innovative methods to examine the extent to which parenting and adjustment are linked at the between- versus within-family level. Compared to methods that do not distinguish between these levels, our approach estimates dynamic reciprocal processes within the family more accurately. Our findings contribute to a growing literature suggesting that stable between-family differences might play a larger role than was previously thought, even in early childhood. Confidence in the results is bolstered by the fact that we used multi-informant reports: Each parent reported on their own parenting, and teachers reported on children's emotional adjustment. This avoids common method bias and single reporter bias. As teachers are likely to be "blind" to parents' childrearing behavior, and have no stake in the parent-child relationship, their reports might be less biased in general.

A potential limitation of the present study is that teachers' reports of emotional adjustment might be more noisy than self- or parent-reports. A teacher is responsible for many pupils, and might not be aware of each pupil's emotional adjustment. Moreover, although the instrument we used assessed *overt* symptoms of emotional maladjustment, people are generally poor judges of others' emotional states [@ickesEmpathicAccuracy1997]. Consequently, our effect sizes might be attenuated by noisy measures of emotional adjustment. Teachers' reports do have clear benefits as well: Young children might lack the ability to report accurately on their own emotional adjustment, and using clinical assessments might be unfeasible for large-scale panel studies. Another limitation to consider is the fact that we operationalized qualitative parenting in terms of specific parenting behaviors, whereas others have focused on parenting *styles* composed of specific combinations of these behaviors [@maccobySocializationContextFamily1983]. We focused on specific parenting behaviors for two reasons. First, although aggregating specific parenting behaviors into parenting styles is theoretically defensible, it also amounts to data reduction. This would prevent a fine-grained investigation of differences between mothers' and fathers' unique roles in childrearing. Moreover, the longitudinal models we used are not suitable for categorical variables, such as parenting style. Still, we have to acknowledge the possibility that the constellation of parenting behaviors is more important than each dimension in isolation. Finally, one potential limitation is that we only included quantitative involvement in the first year of the study. This was done for pragmatic reasons: Children's developmental needs change rapidly, especially in the first years of life. It would be impossible to establish measurement invariance for quantitative involvement from age 6 to 12. Our approach assumes that a "snapshot" of quantitative involvement at age 6 taps into parents' general level of involvement. Consequently, we are unable to assess whether within-person fluctuations in qualitative and quantitative involvement interact. Future research intending to examine such interactions at the within-family level might instead choose to focus on a more restricted age range.

>>>>

abstract: |
  The question whether fathers and mothers play similar roles in children’s
  emotional adjustment has been approached differently in different research
  traditions: Developmental psychologists have focused on qualitative parenting,
  and sociologists have focused on quantitative child care involvement. This
  longitudinal study (N = 1884, age 6-12) bridges these approaches, 
  investigating qualitative and quantitative involvement, and their interaction,
  in relation to emotional adjustment. On average, evidence for mothers'
  and fathers' unique roles was limited.
  When both parents were more involved, both engaged
  more in desirable qualitative parenting. When mothers were disproportionately
  involved, however, parenting roles were more differentiated:
  Fathers displayed less
  desirable qualitative parenting, and children were less well-adjusted.
  This suggests that all family members benefit from a more equal
  childcare distribution.

@@

\vspace{\baselineskip}
Emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001]. It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013]. Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of the quantitative amount of care parents provide (who does what), and in terms of qualitative parenting behaviors (how do parents behave towards children). Nevertheless, the existing literature has two major shortcomings. Firstly, quantitative involvement and qualitative parenting behaviors are rarely jointly investigated. Therefore, we currently don't know how the interplay between quantitative and qualitative involvement relates to emotional development. Secondly, although fathers have been theorized to play a unique role in emotional adjustment development, empirical work has rarely addressed the joint and relative contributions of fathers' and mothers' involvement. The present study addresses overcomes these limitations by incorporating both quantitative and quantitative measures of mothers' and fathers' involvement. We use sophisticated methods to examine how joint and relative quantitative involvement moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment. We thus obtain a more comprehensive and nuanced understanding of how fathers' and mothers' involvement relate to children's emotional adjustment.

@@

Attachment relationships are grounded in a balance between security and exploration [@grossmannDeveloppementAttachementAdaptation1998]. Fatherhood researchers have posited that warmth and security are more central to mother-child attachment relationships, whereas father-child relationships have been construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013]. By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000]. Consequently, fathers and mothers are theorized influence emotional adjustment development in somewhat different ways.

The past decades have seen major strides in the consolidation of theories to guide fatherhood research. Early fatherhood research primarily operationalized involvement in *quantitative* terms: Presence versus absence in divorced families, and the sharing of childcare duties in dual-earner families [@lambBiosocialPerspectivePaternal1987]. Over time, the notion the fathers should share child-rearing responsibilities with mothers, adopting roles historically fulfilled by mothers, gained ground [@pleckFatherhoodIdealsUnited1997]. Consequently, definitions of father involvement evolved to incorporate *qualitative* dimensions, such as warmth and control [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].

It is now widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004]. Nevertheless, these dimensions are rarely investigated together [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014]. The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.

@@

The literature linking demandingness and emotional adjustment is less conclusive [@morrisRoleFamilyContext2007]. Demandingness encompasses several parenting practices related to setting and enforcing expectations for children's behavior. Some of these might help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013]. We examined two aspects of demandingness: Consistency; the extent to which consequences for transgressions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments. Consistency has been found to correlate strongly with emotional adjustment [@brandPerceivedParentalInconsistency1990]. Inductive reasoning is thought to promote the internalization of parental expectations, including display rules for emotions, and might thus play a role in emotional adjustment development as well [@grusecImpactParentalDiscipline1994].

@@

With regard to differential involvement, we might also expect the more involved parent to display more desirable qualitative parenting behaviors, due to the effects of experience [@sabattiniRelationMothersFathers2004]. However, the benefit of parents monitoring each other's child-rearing should be diminished when parents are not equally involved [@mclanahanConsequencesFatherAbsence2014]. Indeed, retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004]. Additionally, greater differential involvement might undermine both parents' involvement quality, through its effect on the marital relationship. Childcare responsibilities are a heavy burden [@deutschPaternalParticipationChild2001]. An unfair distribution strains the interparental relationship, interferes with sensitive parenting, and undermines child well-being [@davies2006interparental]. When care is distributed relatively equally, mothers tend to be more satisfied [@pharesFathersMothersInvolvement2009], and the family system more harmonious. Consequently, balanced co-parenting is associated with greater parental warmth [@barnettInterdependenceParentingMothers2008]. We might thus expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment.

@@

Bioecological theories of development indicate that children require a certain amount of care, and who provides this care is, to some extent, flexible [see @bronfenbrennerEcologyHumanDevelopment1979]. This suggests that the total amount of care provided should be positively associated with children's emotional adjustment. Differential involvement, on the other hand, might be negatively linked to emotional adjustment: When children perceive that one of their parents is relatively uninvolved, they likely feel less supported than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001].

@@

Interactions between involvement quality and quantity have largely been overlooked in theoretical models of father involvement; perhaps in part because these two dimensions have rarely been jointly investigated [@jaffeeLifeFatherBenefits2003]. In this respect, theories of father involvement can be complemented by Feinberg's coparenting model [@feinbergInternalStructureEcological2003]. Coparenting refers to the ways parents relate to each other in the role of parent. Although the coparenting literature is typically invoked to describe child-rearing in divorced families, it also holds insights for intact families. According to Feinberg, the "division of labor" in childcare, analogous to our operationalization of differential involvement, affects qualitative child-rearing behaviors and children's adjustment directly, and moderates the link between these two constructs.

@@

Emotional adjustment was measured using the reverse-coded emotional distress subscale of the teacher-reported Strengths and Difficulties Questionnaire, on a three-point scale from "not true" to "certainly true" [e.g., "Often unhappy, downhearted or tearful"; @goodmanPsychometricPropertiesStrengths2001].

@@

We calculated mean scores for all of the study variables. Scale descriptive statistics are provided in Supplementary Table 1, and scale correlations are provided in Supplementary Table 2. We deemed it defensible to assume measurement invariance, as the difference in fit between structurally invariant models, and metrically invariant models, was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).

To address our research questions, we used multilevel, autoregressive structural equation modeling [@asparouhovDynamicStructuralEquation2018]. In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean). This has several advantages: First, controlling for stable between-family differences allows us to estimate the direction of within-family effects more accurately [@hamakerCritiqueCrosslaggedPanel2015]. Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts. Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects. In this way, we included total joint childcare, the childcare distribution, and child sex as moderators.

@@

With regard to child effects, in line with hypotheses, the intercept of the cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabcrosspath)). Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more anger and inductive reasoning, but also more warmth, than usual in the next year. Contrary to hypotheses, the intercept of the effect of emotional adjustment on maternal consistency was non-significant. As hypothesized, the intercept for maternal inductive reasoning was positive. Finally, counter to hypotheses, the intercept for maternal anger was positive, not negative. In partial support of hypotheses, the intercept for the child effect of adjustment on paternal warmth was positive and significant. None of the other intercepts for child effects on paternal parenting were significant. The variances of all child effects were significant, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs between families.

@@

When both parents were more involved, they displayed greater warmth, consistency, and inductive reasoning. Fathers, in particular, reported substantially greater parental warmth when both parents were more involved. We found that, when mothers performed more care tasks than fathers, mothers displayed more warmth, and fathers displayed less warmth, consistency, and inductive reasoning. This increase in maternal warmth might reflect an attempt to compensate for fathers' absence [@shinnFatherAbsenceChildren1978]. These results are in line with our hypothesis that greater involvement would be associated with more desirable parenting behaviors [@sabattiniRelationMothersFathers2004]. Additionally, they suggest that, when childcare is unequally distributed, fathers engage in less desirable qualitative parenting behaviors [@craigHowMothersFathers2011].

@@

Our operationalization of quantitative involvement, in terms of total joint care and the childcare distribution, allowed us to addresses the claim that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001]. However, counter to our hypothesis, joint care was *negatively* associated with between-family differences in emotional adjustment. One potential explanation is suggested by self-determination theory: Excessive levels of involvement might undermine children's autonomy needs, thereby leading to increased emotional maladjustment [@vandergiessenPerceivedAutonomySupport2014]. Research on "helicopter parenting" supports this interpretation, as psychological well-being was found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014]. Another potential explanation is that the amount of quantitative care provided is not only driven by parents' motivation and ability to be involved in their children's lives, but also by children's objective needs. Factors related to children's objective needs - such as developmental delay, or atypical development - might thus drive an increase in parents' quantitative involvement, and might also be associated with children's emotional maladjustment, thereby resulting in a negative correlation between the two.

In line with predictions derived from coparenting theory, we found that children's emotional adjustment was lower in families where mothers were more involved than fathers - even after controlling for the total amount of care provided. Although substantial empirical work has indicated that the marital relationship suffers when childcare is unequally distributed [@shockleyItNotWhat2018], our research is one of few to examine the implications for children's adjustment [see also @deutschPaternalParticipationChild2001]. These correlational findings at the between-family level should not be interpreted causally, however. Future research should examine potential pathways, such as whether children's adjustment suffers when they feel like both parents are not equally invested in their well-being, or whether fathers are more likely to withdraw from childcare responsibilities when children display poor emotional adjustment, or whether differential involvement strains the marital relationship, and this strain spills over to children's emotional adjustment.

@@

Several strengths allow us to make unique contributions to the literature. First, mothers and fathers both completed identical parenting measures, which allowed us to disentangle and compare their unique roles in children's emotional adjustment development. Second, we incorporated both quantitative and qualitative involvement. This allowed us to provide a more nuanced overview of the links between parental involvement and emotional adjustment development than prior work, including how the distribution of quantitative childcare relates to parenting quality and emotional adjustment, and moderates links between these two constructs. Although some prior studies have also incorporated both quality and quantity of parenting, this work did not disentangle the total amount of care children received from the relative distribution of care tasks [see @deutschPaternalParticipationChild2001]. We addressed this limitation by re-expressing the variance of maternal and paternal quantitative involvement into joint total involvement, and differential involvement. Third, we used innovative methods to examine the extent to which parenting and adjustment are linked at the between- versus within-family level. Compared to methods that do not distinguish between these levels, our approach estimates dynamic reciprocal processes within the family more accurately. Our findings contribute to a growing literature suggesting that stable between-family differences might play a larger role than was previously thought, even in early childhood. Confidence in the results is bolstered by the fact that we used multi-informant reports: Each parent reported on their own parenting, and teachers reported on children's emotional adjustment. This avoids common method bias and single reporter bias. As teachers are likely to be "blind" to parents' child-rearing behavior, and have no stake in the parent-child relationship, their reports might be less biased in general.

A potential limitation of the present study is that teachers' reports of emotional adjustment might be more noisy than self- or parent-reports. A teacher is responsible for many pupils, and might not be aware of each pupil's emotional adjustment. Moreover, although the instrument we used assessed *overt* symptoms of emotional maladjustment, people are generally poor judges of others' emotional states [@ickesEmpathicAccuracy1997]. Consequently, our effect sizes might be attenuated by noisy measures of emotional adjustment. Teachers' reports do have clear benefits as well: Young children might lack the ability to report accurately on their own emotional adjustment, and using clinical assessments might be unfeasible for large-scale panel studies. Another limitation to consider is the fact that we operationalized qualitative parenting in terms of specific parenting behaviors, whereas others have focused on parenting *styles* composed of specific combinations of these behaviors [@maccobySocializationContextFamily1983]. We focused on specific parenting behaviors for two reasons. First, although aggregating specific parenting behaviors into parenting styles is theoretically defensible, it also amounts to data reduction. This would prevent a fine-grained investigation of differences between mothers' and fathers' unique roles in child-rearing. Moreover, the longitudinal models we used are not suitable for categorical variables, such as parenting style. Still, we have to acknowledge the possibility that the constellation of parenting behaviors is more important than each dimension in isolation. Finally, one potential limitation is that we only included quantitative involvement in the first year of the study. This was done for pragmatic reasons: Children's developmental needs change rapidly, especially in the first years of life. It would be impossible to establish measurement invariance for quantitative involvement from age 6 to 12. Our approach assumes that a "snapshot" of quantitative involvement at age 6 taps into parents' general level of involvement. Consequently, we are unable to assess whether within-person fluctuations in qualitative and quantitative involvement interact. Future research intending to examine such interactions at the within-family level might instead choose to focus on a more restricted age range.
